---
title: "How to Work with Spatial Lines and Polygons"
format: html
editor: visual
---

Install and declare required package

```{r}
library("terra")
library("tidyverse")
```

Upload light rail shapefiles into the environment

```{r}
lr <- vect("G:/Shared drives/2024 FIRE-SA/DATA/Silver Lining/zipcodes/tl_2020_us_zcta520.shp")
```

```{r}
lrdf <- as.data.frame(lr)
```

```{r}
#Charlotte
lr<-vect("G:\Shared drives\2024 FIRE-SA\DATA\Silver Lining\Metro_Lines_Regional\Metro_Lines_Regional.shp")
charlotte_lr<-aggregate(lr, dissolve=TRUE)

#Phoenix-Mesa
lr <- vect("Shapefiles/Valley_Metro_Light_Rail/LightRailLine.shp")
lr2 <- subset(lr, lr$SYMBOLOGY=="METRO")
phoenix_mesa_lr<-aggregate(lr2, dissolve=TRUE)

```

## Get PUMA - Example for Phoenix-Mesa, AZ

Upload a PUMA boundary into the Environment

```{r}
puma_az<-vect("Shapefiles/tl_2010_AZ_puma10/tl_2010_04_puma10.shp")
```

Transform the coordinate reference system of the light rail route to be the same as the PUMA boundaries

```{r}
phoenix_mesa_lr_project<-project(phoenix_mesa_lr, crs(puma_az))
```

Map the light rail routes and PUMA boundaries together

```{r}

plot(phoenix_mesa_lr_project, col="blue") #the first item determines the boundary
plot(puma_az, add=TRUE)

```

Find which PUMAs the light rail goes through

```{r}
lr_intersect<-terra::intersect(phoenix_mesa_lr_project, puma_az)

lr_intersect_df<-as.data.frame(lr_intersect) %>%
  mutate(status="LR Overlaps")

print(lr_intersect_df$PUMACE10)
```

Find which PUMAs are around the 1 km buffer of the light rail

```{r}
#make 1 km light rail buffer
lr_buff<-buffer(phoenix_mesa_lr_project, width=1000)

lr_buff_intersect<-terra::intersect(lr_buff, puma_az)

lr_buff_intersect_df<-as.data.frame(lr_buff_intersect) %>%
  mutate(status="Inside 1km Buffer")

print(lr_buff_intersect_df$PUMACE10)
```

Visualize light rail buffer

```{r}
plot(lr_buff, col="blue") #the first item determines the boundary
plot(puma_az, add=TRUE)
```

Combine two dataframes

```{r}
phoenix_mesa_puma<-rbind(lr_intersect_df, lr_buff_intersect_df)
```
