---
title: "Preliminary Research Results"
author: "2024 FIRE198 Sustainability Analytics"
format: gfm
editor: visual
---

## Team Members

Write names of team members here.

Raphael Coronel, Taylor Woods, Brian Huang, Joseph Tang

## Research Question

Write research question here.

How does the opening/announcement of the Silver Line affect housing costs?

## Background

Write research relevance and literature review here.

## Data

Explain data source here.

Clean data here.

```{r}
library("tidyverse")
library("glue")
```

## Preliminary Results

Use your final dataset to visualize how the treatment variable impacts the outcome variable here.

First I add a column saying how many bathrooms the datapoint has has. Only code for br5 is shown but imagine the samw thing happens fo rthe other 4.

```{r}
br5$BedroomCnt <- rep(5,nrow(br5))
```

creating one big data frame and double checking that nothing got nipped off and moving bedroomcnt to the front since its kinda imortant

```{r}
brAll <- rbind(br1,br2,br3,br4,br5)
sum(nrow(br1),nrow(br2),nrow(br3),nrow(br4),nrow(br5))
br <- brAll %>% relocate(BedroomCnt)
```

Some additional processing to get rid of horrendous date data and replace with good year + month data 😌

```{r}
br2 <- br  %>% pivot_longer(cols = `2000-01-31`:`2024-07-31`,names_to = "Date", values_to = "Price")
br3 <- br2 %>% mutate(Year = year(Date), Month = month(Date))
br3$Date <- NULL
```

savin it for now

```{r}
write.csv(br3,file = "Zhvi.csv")
```

due to git file size constraints, the file is saved at:

### G:\Shared drives\\2024 FIRE-SA\DATA\Silver Lining
