
---
title: "Preliminary Research Results"
author: "2024 FIRE198 Sustainability Analytics"
format: gfm
editor: visual
---

## Team Members

Write names of team members here.

Raphael Coronel, Taylor Woods, Brian Huang, Joseph Tang

## Research Question

Write research question here.

How does the opening/announcement of the Silver Line affect housing costs?

## Background

Write research relevance and literature review here.

## Data

Explain data source here.

Clean data here.

```{r}
library("tidyverse")
library("glue")
```

## Preliminary Results

Use your final dataset to visualize how the treatment variable impacts the outcome variable here.

First I add a column saying how many bathrooms the datapoint has has. Only code for br5 is shown but imagine the samw thing happens fo rthe other 4.

```{r}
br5$BedroomCnt <- rep(5,nrow(br5))
```

creating one big data frame and double checking that nothing got nipped off and moving bedroomcnt to the front since its kinda imortant

```{r}
brAll <- rbind(br1,br2,br3,br4,br5)
sum(nrow(br1),nrow(br2),nrow(br3),nrow(br4),nrow(br5))
br <- brAll %>% relocate(BedroomCnt)
```

Some additional processing to get rid of horrendous date data and replace with good year + month data 😌

```{r}
br2 <- br  %>% pivot_longer(cols = `2000-01-31`:`2024-07-31`,names_to = "Date", values_to = "Price")
br3 <- br2 %>% mutate(Year = year(Date), Month = month(Date))
br3$Date <- NULL
```

savin it for now

```{r}
write.csv(br3,file = "Zhvi.csv")
```

due to git file size constraints, the file is saved at:

### G:\Shared drives\\2024 FIRE-SA\DATA\Silver Lining
=======
---
title: "Research Question"
format: gfm
editor: visual
---

Step 1. Install necessary packages.

```{r}
install.packages("tidyverse")
install.packages("kableExtra")
```

Step 2. Declare that you will use these packages in this session.

```{r}
library("tidyverse")
library("kableExtra")
```

Step 3. Upload the dataframe that you have created in Spring 2024 into the repository.

Step 4. Open the dataframe into the RStudio Environment.

```{r}
df<-read.csv("data.csv")
```

Step 5. Use the **head** and **kable** function showcase the first 10 rows of the dataframe to the reader.

```{r}
kable(head(df))
```

## Question 1: What is the frequency of this data frame?

Answer:

## Question 2: What is the cross-sectional (geographical) unit of this data frame?

Answer:

Step 6. Use the **names** function to display all the variables (column) in the dataframe.

```{r}
names(df)

```

## Question 3: Which column represents the treatment variable of interest?

Answer:

## Question 4: Which column represents the outcome variable of interest?

Answer:

Step 7: Create a boxplot to visualize the distribution of the outcome variable under treatment and no treatment.

```{r}
ggplot(df, aes(x=nox_emit)) +
  geom_histogram() +
  facet_wrap(~nbp)
```

Step 8: Fit a regression model $y=\beta_0 + \beta_1 x + \epsilon$ where $y$ is the outcome variable and $x$ is the treatment variable. Use the **summary** function to display the results.

```{r}
model1<-lm(nox_emit ~ nbp, data=df)

summary(model1)
```

## Question 5: What is the predicted value of the outcome variable when treatment=0?

Answer:

## Question 6: What is predicted value of the outcome variable when treatment=1?

Answer:

## Question 7: What is the equation that describes the linear regression above? Please include an explanation of the variables and subscripts.

Answer:

## Question 8: What fixed effects can be included in the regression? What does each fixed effects control for? Please include a new equation that incorporates the fixed effects.

Answer:

## Question 9: What is the impact of the treatment effect once fixed effects are included?

Answer:

# Questions for Week 5

## Question 10: In a difference-in-differences (DiD) model, what is the treatment GROUP?

Answer:

## Question 11: In a DiD model, what are the control groups?

Answer:

## Question 12: What is the DiD regression equation that will answer your research question?

## Question 13: Run your DiD regressions below. What are the results of the DiD regression?

## Question 14: What are the next steps of your research?

Step 9: Change the document format to gfm

Step 10: Save this document as README.qmd

Step 11: Render the document. README.md file should be created after this process.

Step 12: Push the document back to GitHub and observe your beautiful document in your repository!

Step 13: If your team has a complete dataframe that includes both the treated and outcome variable, you are done with the assignment. If not, make a research plan in Notion to collect data on the outcome and treatment variable and combine it into one dataframe.


Step 13: If your team has a complete dataframe that includes both the treated and outcome variable, you are done with the assignment. If not, make a research plan in Notion to collect data on the outcome and treatment variable and combine it into one dataframe.

